% !TEX root = main.tex
%en preliminär problemformulering satt i relation till litteraturbasen
\chapter{Introduction}\label{cha:intro}

\section{Background}
High precision positioning systems are vital in e.g. scanning tunneling microscopes (\abbrSTM), atomic force microscopes (\abbrAFM) and in semiconductor lithography. In \abbrAFM, for instance, high precision positioning is required to control the vertical position of the scanning probe to keep the force constant between the sample surface and the probe tip. An topographical image of the sample is obtained by raster-scanning the probe over the sample surface and plotting the vertical displacement against the probe's x-y position. A positioning system that keeps the force constant down to an atomic-scale resolution is thus inevitable in order to obtain a high resolution image without damaging the sample \citep{SurveyOfControlIssues:2007}.

The piezoelectric effect is a phenomenon that arises in certain solid materials when an electric potential is generated in response to applied mechanical stress. The effect was first discovered by Jacques and Pierre Curie in 1880 when they found that applying pressure to a quarz crystal generates electrical potential. Today, the effect is commonly encountered in daily life and utilized in for example lighters, buzzers and loudspeakers.

Smart materials such as piezoelectric and magnetostrictive materials are nowadays commonly used in precision actuators due to their ability to convert electrical energy into mechanical energy. Piezoelectric materials have been commercially available for almost 45 years and have become indispensable for the nanopositioning industry \citep{Piezo:2008}. In cases where a relatively small displacement range is required (travel ranges up to \unit{500}{\micro\meter}) a piezo electric device is the actuator of choice due to its fast response, high resolution and its ability to generate large mechanical forces for small amounts of power in compact designs \citep{SurveyOfControlIssues:2007}.

The Equipment Controls and Electronics section in the Engineering Department at \abbrCERN (European Organization for Nuclear Research) is developing a high precision positioning system for the control of a piezo-actuated rotational stage used in the UA9 Crystal Collimation study. The stage uses a piezo electric linear stack actuator to displace a flexible lever arm mechanism which generates the rotational movement.

\section{Purpose and Goal}
Crystalline solids have the ability to constrain the directions that particles take as they pass through, this is commonly called the "channelling" property. The UA9 collaboration at \abbrCERN is investigating how tiny bent crystals can help to steer particle beams in modern hadron colliders such as the Large Hadron Collider (\abbrLHC) \citep{WebsiteUA9:2016}. In high energy colliders, such as the \abbrLHC, particles tends to drift outwards creating a beam halo. These particles surrounding the beam, can be lost and cause damage to sensitive parts in the accelerator. By using bent crystals, halo particles can be efficiently extracted from the beam and collected by absorbers further away, reducing the complexity of the system. One major difficulty that arises is that the higher the energy of the particle, the lower the angular acceptance for channeling. Hence, a high precision positioning mechanism with a high angular accuracy is required. The rotational stage (with a range of \unit{20}{\milli\rad}) is of necessity to be able to track reference trajectories at ramp rates of \unit{100}{\micro\radianpersecond} and reject external disturbances (e.g. from the linear axis movement) to maintain a maximum tracking error of $\pm$\unit{1}{\micro\rad}.

This project aims to identify the possible control approaches that could be applicable to this problem to achieve the desired performance.

\section{Prospective challenges}
First of all, piezoelectric  actuators show strong nonlinear properties such as hysteresis and creep (drift), which have to be compensated for. Moreover, the mechanical flexural structure in combination with the piezo electric characteristics leads to a highly resonant structure, making it difficult to achieve the desired performance while operating the rotational stage within noisy environments with external disturbances such as ground vibrations. Furthermore this rotational stage is mounted on a linear stage which is composed of a leadscrew, a stepping motor and a gearbox. The linear movement adds additional perturbations to the rotational stage due to imperfections in the leadscrew and detent torque and stepping nature of the motor. Finally the system dynamics also show linear position and voltage level dependence requiring a controller that is robust to such variations.

One attempt to achieve the desired performance has already been made. The proposed controller, presented in \citep{ButcherController:2015} delivers reasonable performance but does not fulfill the requirements during linear axis movement. The authors proposes a PID controller in combination with a lead-, notch- and prefilter. The controller also uses a hysteresis compensator to mitigate the nonlinear behaviour. The controller has shown high disturbance rejection at the first resonance peak as well as good tracking performance.

\section{Outline}
This thesis plan presents an overview of the thesis, including method, literature base and expected results. A literature review is presented in  Chapter~\ref{cha:litreview}. The method and work flow of this thesis as well as a brief timeplan is given in Chapter~\ref{cha:method}. In Chapter~\ref{cha:result} the results that can be expected half way through the project are discussed while a brief summary of the thesis can be found in Chapter~\ref{cha:conclusion}.
